### 2014-05-02 prevpub.tex
Include merged bibliography

\chapter*{Previously Published Material}
\addcontentsline{toc}{chapter}{Previously Published Material}%

{
\setlength{\parindent}{0 pt}
\setlength{\parskip}{\baselineskip}

Chapter~\ref{chap:sprout}~revises a previous
publication~\cite{sprout}: Keith Winstein, Anirudh Sivaraman, and Hari
Balakrishnan.  \newblock {Stochastic Forecasts Achieve High Throughput
  and Low Delay over Cellular Networks}.  \newblock In {\em NSDI},
2013.

Chapter~\ref{chap:remy}~revises a previous publication~\cite{remy}:
Keith Winstein and Hari Balakrishnan.  \newblock {TCP ex Machina:
  Computer-Generated Congestion Control}.  \newblock In {\em SIGCOMM},
2013.

Chapter~\ref{chap:learnability}~is adapted from a conference
submission that will appear as: Anirudh Sivaraman, Keith Winstein,
Pratiksha Thaker, and Hari Balakrishnan. \newblock {An Experimental
  Study of the Learnability of Congestion Control}. \newblock In {\em
  SIGCOMM}, 2014.

Chapter~\ref{chap:mosh}~revises a previous publication~\cite{mosh}:
Keith Winstein and Hari Balakrishnan.  \newblock {Mosh: An Interactive
  Remote Shell for Mobile Clients}.  \newblock In {\em USENIX Annual
  Technical Conference}, 2012.  }