### 2014-05-03 dotgraphs3/graphs.tex
Import Sprout paper

\documentclass{article}

\usepackage{fullpage, graphicx, times}

\usepackage[usenames, dvipsnames]{xcolor}

\begin{document}
\pagestyle{empty}
\thispagestyle{empty}
\footnotesize
\noindent \begin{tabular}{lp{-0 cm}r}
\def\svgwidth{3.25 in}\input{VerizonLTE-Downlink.pdf_tex} & &
\def\svgwidth{3.25 in}\input{VerizonLTE-Uplink.pdf_tex} \\
\\
\def\svgwidth{3.25 in}\input{Verizon3G1xEV-DO-Downlink.pdf_tex} & &
\def\svgwidth{3.25 in}\input{Verizon3G1xEV-DO-Uplink.pdf_tex} \\
\\
\def\svgwidth{3.25 in}\input{ATTLTE-Downlink.pdf_tex} & &
\def\svgwidth{3.25 in}\input{ATTLTE-Uplink.pdf_tex} \\
\\
\def\svgwidth{3.25 in}\input{T-Mobile3GUMTS-Downlink.pdf_tex} & &
\def\svgwidth{3.25 in}\input{T-Mobile3GUMTS-Uplink.pdf_tex} \\

\end{tabular}

\end{document}