### 2014-05-03 sprout_acks.tex

\section{Acknowledgments}
\label{s:acks}

We thank Shuo Deng, Jonathan Perry, Katrina LaCurts, Andrew McGregor,
Tim Shepard, Dave T\"{a}ht, Michael Welzl, Hannes Tschofenig, and the
anonymous reviewers for helpful discussion and feedback. This work was
supported in part by NSF grant 1040072. KW was supported by the Claude
E.~Shannon Research Assistantship. We thank the members of the MIT
Center for Wireless Networks and Mobile Computing, including
Amazon.com, Cisco, Google, Intel, Mediatek, Microsoft, ST
Microelectronics, and Telefonica, for their support.