### 2014-05-03 sprout_problem.tex

%\section{Cellular Wireless Videoconferencing: Context and Challenges}
\section{Context and Challenges}
\label{s:problem}

This section highlights the networking challenges in designing an
adaptive transport protocol on cellular wireless networks. We discuss
the queueing and scheduling mechanisms used in existing networks,
present measurements of throughput and delay to illustrate the
problems, and list the challenges.

\subsection{Cellular Networks}

At the link layer of a cellular wireless network, each device (user)
experiences a different time-varying bit rate because of variations in
the wireless channel; these variations are often exacerbated because of
mobility. Bit rates are also usually different in the two directions
of a link. One direction may experience an outage for a few seconds
even when the other is functioning well. Variable link-layer bit rates
cause the data rates at the transport layer to vary. In addition, as
in other data networks, cross traffic caused by the arrival and
departure of other users and their demands adds to the rate
variability.

Most (in fact, all, to our knowledge) deployed cellular wireless
networks enqueue each user's traffic in a separate queue. The base
station schedules data transmissions taking both per-user
(proportional) fairness and channel quality into
consideration~\cite{propfair}. Typically, each user's device is
scheduled for a fixed time slice over which a variable number of
payload bits may be sent, depending on the channel conditions, and
users are scheduled in roughly round-robin fashion. The isolation
between users' queues means that the dominant factor in the end-to-end
delay experienced by a user's packets is {\em self-interaction},
rather than cross traffic. If a user were to combine a high-throughput
transfer and a delay-sensitive transfer, the commingling of these
packets in the same queue would cause them to experience the same
delay distributions. The impact of other users on delay is
muted. However, competing demand can affect the throughput that a user
receives.

%The impact of other users is not as significant,
%because the time scales over which users arrive and leave is large
%compared to the time scales of channel variations.

Many cellular networks employ a non-trivial amount of packet
buffering. For TCP congestion control with a small degree of
statistical multiplexing, a good rule-of-thumb is that the buffering
should not exceed the bandwidth-delay product of the connection. For
cellular networks where the ``bandwidth'' may vary by two orders of
magnitude within seconds, this guideline is not particularly useful. A
``bufferbloated''~\cite{bufferbloat} base station at one link rate
may, within a short amount of time, be under-provisioned when the link
rate suddenly increases, leading to extremely high IP-layer packet
loss rates (this problem is observed in one
provider~\cite{Mahajan12}).

The high delays in cellular wireless networks cannot simply be blamed
on bufferbloat, because there is no single buffer size that will
always work. It is also not simply a question of using an appropriate
Active Queue Management (AQM) scheme, because the difficulties in
picking appropriate parameters are well-documented and become harder
when the available rates change quickly, and such a scheme must be
appropriate when applied to all applications, even if they desire bulk
throughput. In \S\ref{s:eval}, we evaluate CoDel~\cite{CoDel}, a
recent AQM technique, together with a modern TCP variant (Cubic, which
is the default in Linux), finding that on more than half of our tested
network paths, CoDel slows down a bulk TCP transfer that has the link
to itself.
%CoDel dramatically improved TCP's
%end-to-end delays compared with a gateway that indefinitely queues
%packets, but not as much as Sprout (even without active queue
%management).
%
%Sprout outperformed Cubic-over-CoDel in delay (with a 60\% reduction)
%but also had 30\% less throughput. Sprout-EWMA surpassed
%Cubic-over-CoDel in throughput (by 31\%) and nearly matched it in
%delay.

By making changes---when possible---at endpoints
instead of inside the network, diverse applications may have more
freedom to choose their desired compromise between throughput and
delay, compared with an AQM scheme that is applied uniformly to all
flows.

%When rate variations occur both because of congestion and because of
%channel variations, a different approach is required to provide high
%throughput at low delay. For applications concerned only with bulk
%throughput (e.g., a file download), TCP running over existing cellular
%networks generally keeps a varying link occupied and achieves close to
%the maximum possible throughput. But this comes at the expense of
%extremely long delays. Installing CoDel at the bottleneck link
%dramatically improves the delay, but hurts throughput. 
%By contrast,
%Sprout provides an end-to-end way to achieve both high throughput and
%low delay---on some networks, higher throughput and lower delay than
%TCP-over-CoDel.

Sprout is not a traditional congestion-control scheme, in that its
focus is directed at adapting to varying link conditions, not to
varying cross traffic that contends for the same bottleneck queue. Its
improvements over existing schemes are found when queueing delays are
imposed by one user's traffic. This is typically the case
when the application is running on a mobile device, because cellular
network operators generally maintain a separate queue for each
customer, and the wireless link is typically the bottleneck.  An
important limitation of this approach is that in cases where these
conditions don't hold, Sprout's traffic will experience the same
delays as other flows.
%We believe this is an unavoidable consequence
%of an end-to-end approach.

\subsection{Measurement Example}

In our measurements, we recorded large swings in available throughput
on mobile cellular links. Existing interactive transports do
not handle these well. Figure~\ref{fig:skypevssprout} shows an
illustrative section of our trace from the Verizon LTE downlink, whose
capacity varied up and down by almost an order of magnitude within one
second. From 15 to 25 seconds into the plot, and from 43 to 49 seconds,
Skype overshoots the available link capacity, causing large standing
queues that persist for several seconds, and leading to glitches or
reduced interactivity for the users. By contrast, Sprout works to
maximize the available throughput, while limiting the risk that any
packet will wait in queue for more than 100~ms (dotted line). It also
makes mistakes (e.g., it overshoots at $t=43$ seconds), but then
repairs them.

Network behavior like the above has motivated our development of
Sprout and our efforts to deal explicitly with the uncertainty of
future link speed variations.

\subsection{Challenges}

A good transport protocol for cellular wireless networks
must overcome the following challenges:
\begin{enumerate}
\item It must cope with dramatic temporal variations in link rates.
\item It must avoid over-buffering and incurring high delays, but at
  the same time, if the rate were to increase, avoid
  under-utilization.
\item It must be able to handle outages without over-buffering, cope
  with asymmetric outages, and recover gracefully afterwards.
%\item At the application layer, the quality of delivered video and
%  audio must match the achievable link rate to provide the best
%  possible experience to the user.
\end{enumerate}

Our experimental results show that previous work (see
\S\ref{s:related}) does not address these challenges
satisfactorily. These methods are reactive, using packet
losses, round-trip delays, and in some cases, one-way delays as the
``signal'' of how well the network is doing. In contrast, Sprout uses
a different signal, the observed arrival times of packets at the
receiver, over which it runs an inference procedure to make
forecasts of future rates. We find that this method produces a good
balance between throughput and delay under a wide range of
conditions.