### 2014-05-03 remy-rebuttal.txt
Import Remy

We thank the reviewers for their detailed and thoughtful feedback.

Reviewer A: We agree that the sec-5.5 results on these RemyCCs when
run concurrently vs. CUBIC/Compound aren't great, but emphasize that
they weren't yet trained to compete with "buffer-filling" TCPs. When
we train against possible buffer-filling cross-traffic, we expect to
demonstrate throughput gains. Please note: CUBIC isn't fair to Newreno
or Compound either.

Reviewer B: We agree with the suggestion to cite Chiu/Jain, who found
that among linear schemes where flows start at the same time with
uniform RTT and last forever, AIMD converges to fairness. RemyCCs
aren't necessarily linear. Beyond [42], we don't know of a theoretical
treatment with these assumptions removed, but our work can give
feasibility results.

We also seek intuition about why RemyCCs do well. The short answer is:
they can switch between pacing and window-based control, and respond
quickly to entry/exit of flows.  We can now trace executing RemyCCs
and could expand the explanation.

Reviewer C:

1. Remy is intended solely for *offline* creation of
congestion-control schemes, which then run at the sender (top-right of
pg. 2; fig. 1; sec. 4.3). Connections start immediately.

2. Remy doesn't assume network paths are fixed. We evaluated on traces
of time-varying Verizon and AT&T LTE networks (figs. 6-8), with actual
flow lengths (fig. 3). Only in *offline training* did we use a mean
on/off time of 5 seconds. (Training parameters are on pg. 7, right
column; evaluation scenarios are under "Topologies.") Other reported
simulation results varied cross traffic, RTT, etc.

3. The network model used for training captures a 64-fold *range* of
BDP-per-user. We evaluated on networks inside and outside this range
(pg. 7, col. 2).

4. Re: XCP, it is *XCP* that needs to know the exact throughput of the
bottleneck link (footnote 8). Remy does not.

5. Remy doesn't assume the exact traffic model is provided; we trained
on a model but evaluated on a real (untrained) flow-length
distribution.

6. We don't claim Remy is optimal; there is an optimal scheme in
principle (sec 4.1), which we attempt to approximate. Section 3.3
stresses that the utility function is a tool used to design
protocols. It need not capture users' "true" utilities (cf. remarks in
Related Work re: stochastic NUM [42]).

7. The evaluation looked at more than just queueing latency. Our
tables (sec. 1) and charts (figs. 4-7,9) show delay-vs-throughput and
delay-fairness and throughput-fairness (shaded ovals).