### 2014-05-03 remy_acks.tex

\section{Acknowledgments}

We are grateful to Anirudh Sivaraman for several contributions to the
simulator and for helpful discussions. We thank Leslie Kaelbling,
Christopher Amato, Scott Shenker, and our shepherd, Ranjita
Bhagwan. We thank Frans Kaashoek and Nickolai Zeldovich for the use of
multicore machines at MIT. KW was supported by the Claude E.~Shannon Research
Assistantship. We thank the members of the MIT Center for Wireless
Networks and Mobile Computing (Wireless@MIT), including Amazon.com,
Cisco, Google, Intel, Mediatek, Microsoft, ST Microelectronics, and
Telefonica, for their support. This work was also supported in part by
NSF grant CNS-1040072.