### 2014-05-03 remy_limits.tex

\section{Discussion}

Much remains unknown about the capabilities and limits of
computer-generated algorithms, much less decentralized algorithms that
cooperate indirectly across a network to achieve a common
goal. Although the RemyCCs appear to work well on networks whose
parameters fall within or near the limits of what they were prepared
for --- even beating in-network schemes at their own game and even
when the design range spans an order of magnitude variation in network
parameters --- we do not yet understand clearly \emph{why} they work,
other than the observation that they seem to optimize their intended
objective well.

We have attempted to make algorithms ourselves that surpass the
generated RemyCCs, without success. That suggests to us that Remy may
have accomplished something substantive. But digging through the
dozens of rules in a RemyCC and figuring out their purpose and
function is a challenging job in reverse-engineering. RemyCCs
designed for broader classes of networks will likely be even
more complex, compounding the problem.

Our approach \emph{increases} endpoint complexity in order to
\emph{reduce} the complexity of overall network behavior. Traditional
TCP congestion control specifies simpler behavior for each endpoint,
but the resulting emergent behavior of a multiuser network is not
easily specified and is often suboptimal and variable, and even
unstable.

By contrast, our approach focuses on maximizing a well-specified
overall objective at the cost of complex endpoint algorithms. We think
this tradeoff is advisable: today's endpoints can execute complex
algorithms almost as easily as simple ones (and with Remy, the bulk of
the intelligence is computed offline). What users and system designers
ultimately care about, we believe, is the quality and consistency of
overall behavior.

Our {\em synthesis-by-simulation} approach also makes it easier to
\emph{discuss} competing proposals for congestion control. Today, it
is not easy to say why one flavor of TCP or tweak may be preferred
over another. But if two computer-generated algorithms differ, there is
a reason: either they make different assumptions about the expected
networks they will encounter, or they have different goals in mind, or
one is better optimized than the other. This formulation allows the
implementer to choose rationally among competing options.

All that said, we have much to learn before computer-generated
algorithms will have proven themselves trustworthy:

\begin{itemize}

\item Other than by exhaustive testing, we don't know how to predict
  the robustness of RemyCCs to unexpected inputs. Do they break
  catastrophically in such situations?

\item How would a RemyCC designed for a 10,000-fold range of
  throughputs and RTTs perform?
%, compared with existing algorithms?
%We do not know how a RemyCC's performance will be affected as
%  its prior knowledge becomes less specific and the optimization
%  procedure more arduous.

\item Although we are somewhat robust against a RemyCC's latching on
  to the peculiarities of a simulator implementation (because RemyCCs
  are designed within Remy but then evaluated within ns-2), we can't
  be certain how well RemyCCs will perform on real networks without
  trying them.

% We have not exhaustively cataloged the behavior of
%even the RemyCCs discussed in this paper.
%, and we have not yet tried to
%design a congestion-control protocol that resists being ``squeezed
%out'' by heterogenous incumbent protocols.

\end{itemize}

We believe that making congestion control a \emph{function} of the
desired ends, and the assumptions we make about the network, is the
solution to allow the Internet and its subnetworks to evolve without
tiptoeing around TCP's assumptions about how networks behave. But many
dots need to be connected before the the Internet at large --- as
opposed to internal networks --- might agree on a model that could be
used to prepare a ``one-size-fits-all'' RemyCC.