### 2014-05-03 remy_model.tex

\section{Modeling the Congestion-Control Problem}
\label{s:model}

We treat congestion control as a problem of distributed
decision-making under uncertainty. Each endpoint that has pending data
must decide for itself at every instant: send a packet, or don't
send a packet.

If all nodes knew in advance the network topology and capacity, and the
schedule of each node's present and future offered load, such
decisions could in principle be made perfectly, to achieve a desired
allocation of throughput on shared links.

In practice, however, endpoints receive observations that only hint at
this information. These include feedback from receivers concerning the
timing of packets that arrived and detection of packets that didn't, and
sometimes signals, such as ECN marks, from within the network itself.
Nodes then make sending decisions based on this partial information
about the network.

Our approach hinges on being able to evaluate quantitatively the merit
of any particular congestion control algorithm, and search for the
best algorithm for a given network model and objective function. We
discuss here our models of the network and cross traffic, and how we
ultimately calculate a figure of merit for an arbitrary congestion
control algorithm.

\subsection{Expressing prior assumptions about the network}

From a node's perspective, we treat the network as having been drawn
from a stochastic generative process. We assume the network is
Markovian, meaning that it is described by some state (e.g.~the
packets in each queue) and its future evolution will depend only on
the current state.

Currently, we typically parametrize networks on three axes: the speed
of bottleneck links, the propagation delay of the network paths, and
the degree of multiplexing, i.e., the number of senders contending for
each bottleneck link. We assume that senders have no control over the
paths taken by their packets to the receiver.

Depending on the range of networks over which the protocol is intended
to be used, a node may have more or less uncertainty about the
network's key parameters. For example, in a data center, the topology,
link speeds, and minimum round-trip times may be known in advance, but
the degree of multiplexing could vary over a large range. A virtual
private network between ``clouds'' may have more uncertainty about the
link speed. A wireless network path may experience less multiplexing,
but a large range of transmission rates and round-trip times.

As one might expect, we have observed a tradeoff between generality
and performance; a protocol designed for a broad range of networks may
be beaten by a protocol that has been supplied with more specific and
accurate prior knowledge. Our approach allows protocol designers to
measure this tradeoff and choose an appropriate design range for their
applications.

\subsection{Traffic model}

Remy models the offered load as a stochastic process that switches
unicast flows between sender-receivers pairs on or off. In a simple
model, each endpoint has traffic independent of the other
endpoints. The sender is ``off'' for some number of seconds, drawn
from an exponential distribution. Then it switches on for some number
of bytes to be transmitted, drawn from an empirical distribution of
flow sizes or a closed-form distribution (e.g.~heavy-tailed
Pareto). While ``on,'' we assume that the sender will not stall until
it completes its transfer.

In traffic models characteristic of data center usage, the off-to-on
switches of contending flows may cluster near one another in time,
leading to incast. We also model the case where senders are ``on'' for
some amount of time (as opposed to bytes) and seek maximum throughput,
as in the case of videoconferences or similar real-time traffic.

\subsection{Objective function}

Resource-allocation theories of congestion control have traditionally
employed the alpha-fairness metric to evaluate allocations of
throughput on shared links~\cite{Srikant}. A flow that receives
steady-state throughput of $x$ is assigned a score of $U_\alpha(x) =
\frac{x^{1-\alpha}}{1-\alpha}$. As $\alpha \rightarrow 1$, in the
limit $U_1(x)$ becomes $\log x$.

Because $U_\alpha(x)$ is concave for $\alpha > 0$ and monotonically
increasing, an allocation that maximizes the total score will prefer
to divide the throughput of a bottleneck link equally between
flows. When this is impossible, the parameter $\alpha$ sets the
tradeoff between fairness and efficiency. For example, $\alpha = 0$
assigns no value to fairness and simply measures total
throughput. $\alpha = 1$ is known as proportional fairness, because it
will cut one user's allocation in half as long as another user's can
be more than doubled. $\alpha = 2$ corresponds to minimum potential
delay fairness, where the score goes as the negative inverse of
throughput; this metric seeks to minimize the total time of
fixed-length file transfers. As $\alpha \rightarrow \infty$,
maximizing the total $U_\alpha(x)$ achieves max-min fairness, where
all that matters is the minimum resource allocations in bottom-up
order~\cite{Srikant}.

Because the overall score is simply a sum of monotonically increasing
functions of throughput, an algorithm that maximizes this total is
Pareto-efficient for any value of $\alpha$; i.e., the metric will
always prefer an allocation that helps one user and leaves all other
users the same or better. Tan et al.~\cite{Tan09} proved that, subject
to the requirement of Pareto-efficiency, alpha-fairness is \emph{the}
metric that places the greatest emphasis on fairness for a particular
$\alpha$.

Kelly et al.~\cite{Kelly98} and further analyses showed that TCP
approximately maximizes minimum potential delay fairness
asymptotically in steady state, if all losses are congestive and link
speeds are fixed.

We extend this model to cover dynamic traffic and network
conditions. Given a network trace, we calculate the average throughput
$x$ of each flow, defined as the total number of bytes received
divided by the time that the sender was ``on.'' We calculate the
average round-trip delay $y$ of the connection.

The flow's score is then 
\begin{equation}
U_\alpha(\textrm{x}) - \delta \cdot U_\beta(\textrm{y}),
\label{eq:util}
\end{equation}
where $\alpha$ and $\beta$ express the fairness-vs.-efficiency
tradeoffs in throughput and delay, respectively, and $\delta$
expresses the relative importance of delay vs.~throughput.% For the
%rest of this paper, we set $\alpha = \beta = 1$ and simply consider
%the sum of $\log(\textrm{throughput}) - \delta \cdot
%\log(\textrm{delay})$ for each flow.

%We also consider the metric of average flow completion time. This also
%expresses a throughput-vs.-delay tradeoff.

We emphasize that the purpose of the objective function is to supply a
quantitative goal from a protocol-design perspective. It need not
(indeed, does not) precisely represent users' ``true'' preferences or
utilities. In real usage, different users may have different
objectives; a videoconference may not benefit from more throughput, or
some packets may be more important than others. We have not yet
addressed the problem of how to accommodate diverse objectives or how
endpoints might learn about the differing preferences of other
endpoints.

%In summary, this framework provides a unitary metric for evaluating
%the performance of any congestion-control endpoint algorithm (e.g.~TCP
%Cubic, Vegas, or Compound) on a given ensemble of network and traffic
%models. In the next section, we describe the design of the Remy
%program, which attempts to design a congestion-control algorithm to
%optimize this metric.