### 2014-05-04 learn_concl.tex

\section{Conclusion}
\label{s:concl}

In the context of network congestion control, we asked: How faithfully
do protocol designers really need to understand the networks they
design for? What are the important signals that endpoints should
listen to?  How can researchers gain confidence that systems that work
in testing scenarios will also perform adequately on real networks
that are inevitably more complex, or future networks yet to be
developed? Is there a tradeoff between the performance of a protocol
and the breadth of its intended operating range of networks?  What is
the cost of playing fairly with cross-traffic that is governed by
another protocol?

We examined these questions using Remy as a design tool to produce a
tractable attempt at optimal (Tao) protocol under various
conditions. We found only weak evidence of a tradeoff between
operating range and performance, even when operating range was
extended to cover a thousand-fold range of link rates. We found that
it may be acceptable to simplify some characteristics of the
network---such as its topology---when modeling for design
purposes. Some other features, such as the degree of multiplexing and
the aggressiveness of contending endpoints, are important to
capture. These results may provide useful insights to network protocol
designers about what factors are important to model accurately and
what factors may be simplified.