### 2014-05-04 learn_learnability.tex

\section{Related Work}
\label{s: learnability}

This section discusses closely related work on congestion control and
explains how our work has an analogy with theoretical notions of
learnability.

\subsection{Congestion control}

Congestion control over packet-switched networks has been a productive
area of research since the seminal work of the 1980s, including
Ramakrishnan and Jain's DECBit scheme~\cite{decbit} and Jacobson's TCP
Tahoe and Reno algorithms~\cite{Jacobson88}. End-to-end algorithms
typically compute a congestion window or a paced transmission rate
using the stream of acknowledgments (ACKs) arriving from the
receiver. In response to congestion, inferred from packet loss or, in
some cases, rising delays, the sender reduces its window or rate;
conversely, when no congestion is perceived, the sender increases its
window or rate.

In this paper, we use the Remy~\cite{remy} protocol-design tool to
generate end-to-end Tao congestion-control schemes from first
principles. The Remy work showed that such an approach can produce
schemes whose performance is competitive with, and often out-performs,
human-generated schemes, including most varieties of TCP congestion
control, on intended target networks.

By contrast, this paper use Remy program as a tool for understanding
the nature of the problem of protocol design without being able to
fully define the intended target networks. We use the program's output
as a proxy for the ``best possible'' Tao congestion-control protocol
intended for a particular imperfect network model, and then ask how
that protocol performs on a different set of networks that varies from
the model in some respect (topology, link speed, behavior of
contending endpoints, etc.).

For reference, we also compare with the most prevalent
congestion-control protocols in wide use, including Linux's TCP
Cubic~\cite{cubic}, and the less-aggressive NewReno
algorithm~\cite{newreno}. End-to-end congestion control may be
assisted with explicit router participation; we also measure the
CoDel~\cite{CoDel} algorithm running with stochastic fair
queueing~\cite{sfq} for ``controlled delay'' at bottleneck gateways.

\subsection{Learnability}

TCP congestion control was not designed with an explicit objective
function in mind. Kelly et al.~present an interpretation of TCP
congestion-control variants in terms of the implicit goals they
attempt to optimize~\cite{Kelly98}.  This line of work is known as
Network Utility Maximization (NUM); more recent work has modeled
stochastic NUM problems~\cite{stochasticnum}, where flows enter and
leave the network with time.

We extend this problem by examining the difficulty of designing a
network protocol given an \emph{imperfect} model of the network where
it will be deployed, in order to understand the inherent difficulties
of the problem of congestion control.

Formally speaking, designing such a protocol is a problem in
sequential decision-making under uncertain and can be modeled as a
Partially-Observable Markov Decision Process~\cite{pomdp}. In that
context, the purpose of this paper is to ask: how well can a protocol
designer ``learn'' the solution to one POMDP or set of networks and
successfully apply the result (a protocol) to a different set of networks?

In doing so, we draw an explicit analogy to the concept of
``learnability'' employed in machine
learning~\cite{learnable,schapire}. A canonical machine-learning
problem attempts to design a classifier for a large population of data
points, supplied with only a smaller (and possibly skewed) ``training
set'' meant to teach the classifier about the full population of
data. Subsequently, the performance of the resulting classifier is
evaluated in how well it correctly classifies points in a ``testing
set,'' generally drawn from the actual population. Learnability theory
measures the difficulty of inferring an accurate classifier for the
testing set, given a training set.

Just as a classifier design procedure may minimize the error rate or
the width of the margin as a proxy for maximizing predictive
performance on unseen inputs, the Remy tool uses an objective function
in terms of throughput and delay, averaged over the design model, as a
proxy for performance on as-yet-unseen networks.

In our work, we envision a protocol designer working to generate a
congestion-control protocol for a large set of real networks (e.g.,
the Internet), supplied with only an imperfect model of the range of
variation and behavior of those networks. The imperfect model is the
``training scenarios''---a model of the target networks used for design
purposes. The ``testing scenarios'' are drawn from the population of actual
target networks.

Learnability here measures the difficulty of designing an adequate
protocol for a network \emph{model}, and then deploying it on real
networks that cannot be perfectly envisioned at the time of design.