### 2014-05-05 prevpub.tex

\chapter*{Previously Published Material}
\addcontentsline{toc}{chapter}{Previously Published Material}%

{
\setlength{\parindent}{0 pt}
\setlength{\parskip}{\baselineskip}

Chapter~\ref{chap:sprout}~revises a previous
publication~\cite{sprout}:
K.~Winstein, A.~Sivaraman, and H.~Balakrishnan.
\newblock {Stochastic Forecasts Achieve High Throughput and Low Delay over
  Cellular Networks}.
\newblock In {\em NSDI}, Lombard, Ill., April 2013.

Chapter~\ref{chap:remy}~revises a previous publication~\cite{remy}:
K.~Winstein and H.~Balakrishnan.
\newblock {TCP ex Machina: Computer-Generated Congestion Control}.
\newblock In {\em SIGCOMM}, Hong Kong, China, August 2013.

Chapter~\ref{chap:learnability}~is adapted from a conference
submission that will appear as:
A.~Sivaraman, K.~Winstein, P.~Thaker, and H.~Balakrishnan
\newblock {An Experimental
  Study of the Learnability of Congestion Control}. \newblock In {\em
  SIGCOMM}, Chicago, Ill., August 2014.

%Chapter~\ref{chap:mosh}~revises a previous publication~\cite{mosh}:
%K.~Winstein and H.~Balakrishnan.
%\newblock {Mosh: An Interactive Remote Shell for Mobile Clients}.
%\newblock In {\em USENIX Annual Technical Conference}, Boston, Mass., June
%  2012.
}