### 2014-05-07 sprout_limit.tex
Finish Sprout section?

\section{Limitations}

Although Sprout's results are encouraging, there are several
limitations. First, as noted in \S\ref{sprout:problem} and
\S\ref{s:sprout}, an end-to-end system like Sprout cannot control
delays when the bottleneck link also carries aggressive competing
traffic that shares the same queue. If a device uses traditional TCP
outside of Sprout, the incurred queueing delay---seen by Sprout and
every flow---will be substantial.

% We leave to
%future work a better understanding of the {\em need} for active queue
%management to achieve good throughput-delay trade-offs with multiple
%independent contending flows, speculating here that a purely
%end-to-end approach may in fact work.

The accuracy of Sprout's forecasts depends on whether the application
is providing offered load sufficient to saturate the link. For
applications that switch intermittently on and off, or don't desire
high throughput, the transient behavior of Sprout's forecasts
(e.g.~ramp-up time) becomes more important. Our evaluation assumed
that a video coder can meet Sprout's contract by supplying on demand,
with no latency, a newly-coded video frame whose compressed length was
exactly equal to the desired amount. In practice, actual video coders
will not be able to meet this contract perfectly, and the resulting
consequences will need to be evaluated.

We have tested Sprout only in trace-based emulation of eight cellular
links recorded in the Boston area in 2012. Although Sprout's model was
frozen before data were collected and was not ``tuned'' in response to
any particular network, we cannot know how generalizable Sprout's
algorithm is without more real-world testing.

Sprout is not a traditional congestion-control protocol, in that it is
designed to adapt to varying link conditions, not varying cross
traffic. In a cellular link where users have their own queues on the
base station, interactive performance will likely be best when the
user runs bulk and interactive traffic \emph{inside} Sprout
(e.g.~using SproutTunnel), not alongside Sprout.

For the classical problem of multiple independent users sharing the
same bottleneck, Sprout is insufficient to provide good results. In
the rest of this dissertation, we expand on Sprout's approach to
address the ``multi-agent'' congestion-control problem.

%In future work, we are eager to explore different stochastic network
%models, including ones trained on empirical variations in cellular
%link speed, to see whether it is possible to perform much better than
%Sprout if a protocol has more accurate forecasts. We think it will be
%worthwhile to collect enough traces to compile a standardized
%benchmark of cellular link behavior, over which one could evaluate any
%new transport protocol.