### 2014-05-21 intro.tex
Tweaks to intro

\chapter{Introduction}
\label{chap:intro}

\begin{figure}
\caption{The Internet's four-layer model}
\label{f:stack}
\begin{centering}
\includegraphics{layers.pdf}

\end{centering}
\end{figure}

\begin{figure}
\caption{As the Internet has evolved, researchers have created at
  least 40 mechanisms to govern resource allocation on the
  network---both entirely distributed schemes (``end-to-end'') and
  ones that include code running ``in-net.''}
\label{f:march}

\vspace{\baselineskip}

\begin{centering}
\noindent \includegraphics[width=\textwidth]{march2-all.pdf}

\end{centering}
\end{figure}

Over the last 25 years, the Internet has transformed from an academic
networking experiment into a global utility with billions of
users. As part of this transformation, every layer of the Internet
``stack'' has seen dramatic change.

At the link layer, technologies that did not exist 25 years
ago now dominate---including wireless local-area networks (Wi-Fi),
cellular networks, datacenter interconnects, and transoceanic links
with high delay.

One layer up, at the Internet layer, mobility is now
ubiquitous. User devices regularly change their interface IP addresses
as they roam from network to network.

At the top layer of the Internet stack---the application
layer---none of the dominant applications of today existed 25 years
ago, including the World Wide Web and its short-length flows,
progressive-download video applications (e.g., YouTube and Netflix),
and real-time streaming video (e.g., Skype and Facetime).

All of the Internet has had to grapple with this continuous
evolution. How should the protocols of the transport layer,
sitting in the middle of the stack between the application and the
Internet, adapt to changing applications above them and
evolving networks below?

One legitimate answer may be that transport-layer protocols need not
adapt. If the network or applications evolve, and a transport
protocol no longer works adequately alongside then, we can simply stop
using the protocol and design a fresh one that matches the new
circumstances.

This approach broadly characterizes the Internet's extraordinarily
successful path over the last 25 years. Looking at just one
function of the transport layer---congestion control, the job of
dividing up the network's resources among contending
users---researchers have accommodated new applications and network
behaviors by devoting considerable effort to develop new
mechanisms, at least 40 in total so far (Figure \ref{f:march}).

Despite its demonstrable success, this approach can have downsides. In
defining transport-layer protocols by their \emph{mechanisms}---by the
actual behavior of endpoints or gateways that execute the
protocol---we leave implicit the assumptions that the mechanism makes
about the behavior of other layers and the policy that the mechanism
is built to pursue.

For contemporary mechanisms (e.g., TCP Cubic, the current default in
Linux), it's difficult to state the assumptions that the transport
makes about lower layers and to predict when those assumptions would
no longer hold. This presents a challenge for link-layer designers who
wish to design a new networking technology. Because of Cubic's
popularity, it has become a \emph{de facto} requirement that Cubic perform
well over any new network. In effect, this means that the designer of
the new link layer must try to satisfy Cubic's implicit assumptions in
order to achieve adequate performance.

This has led to the ``bufferbloat''\cite{bufferbloat} problem:
believing that the transport layer will interpret losses as a
signal to slow down, the link layer tries to drop packets as
rarely as possible. It has also led to a lack of parallelism in
Internet routing---flows only take one path, even if speedups might be
available by splitting flows across multiple routes---on the grounds
that the transport protocol might interpret out-of-order packets
as a pathology.

On the flip side, it is usually impossible to adjust the transport
layer's assumptions in order to accommodate a new kind of
link. There's no way to tweak the assumptions and then retrace the
same design process that the protocol's designers followed, to find the
mechanism they would have produced given a different starting point.

%\section{Showing our work in protocol design}

This thesis proposes essentially that approach to protocol design. The
transport layer should adapt to whatever the layers below may do, and
whatever the application above it wants done. Protocol designers
should specify the \emph{policy}---namely, what assumptions they want
to make about the network and what kind of performance the application
is interested in---and let computers worry about translating that into
the mechanism of congestion control.

By ``showing our work'' in this way, clearly enough for a computer to
recreate the same design, altering the assumptions is a matter of
changing the inputs to a computer program. This makes it easier for
adjacent layers to evolve. My colleagues and I have also found that
this approach can yield better performance than conventional,
human-designed protocols.

\section{Summary of results}

Over the last three years, my colleagues and I built a series of
systems that demonstrate the benefits of automated, objective-driven
protocol design.

\subsection{Sprout}

\textbf{Sprout} (Chapter~\ref{chap:sprout}) is a transport protocol
designed to carry high-throughput interactive traffic, such as a
videoconference, over a cellular network. Sprout includes an explicit
model of the dynamics of cellular networks and makes predictions about
future performance by inferring, with uncertainty, the current state
of the network and evolving the model forward. Its control
strategy---how much data to send at a given moment---is a function of
those predictions and of an explicit objective: maximize throughput,
but bound in-network delays to be at most 100~ms with high
probability.

In a trace-driven experimental evaluation (details in
\S\ref{sprout:eval}), Sprout gave 2-to-4 times the throughput and
7-to-9 times less delay than Skype, Apple Facetime, and Google
Hangouts:

\begin{center}
\noindent \begin{tabular}{|l|c|c|}
\hline
Sprout vs. & Avg.~speedup & Delay reduction \\
\hline
\hline
Skype & $2.2\times$ & $7.9\times$\\
Hangout & $4.4\times$ & $7.2\times$\\
Facetime & $1.9\times$ & $8.7\times$\\
\hline
Compound & $1.3\times$ & $4.8\times$\\
TCP Vegas & $1.1\times$ & $2.1\times$\\
LEDBAT & no change & $2.8\times$\\
Cubic & \cellcolor{red!20}$0.91\times$ & $79\times$\\
%\hline
%Cubic-CoDel & \cellcolor{red!20}$0.70\times$ & $1.6\times$ (0.50~s) \\
%CUBIC/CoDel & & \\
%Compound/CoDel & & \\
\hline
\end{tabular}

{\footnotesize Adapted from Figure~\ref{f:sproutcompe2e}.}

\end{center}

\subsection{Remy}

\textbf{Remy} (Chapter~\ref{chap:remy}) generalizes Sprout to address
the classical problem of \emph{multi-agent} congestion control, where
independent users contend for the same limited network resource. Remy
is a protocol-design tool that takes, as input, a set of assumptions
about the uncertain network and workload, and an objective to pursue
on behalf of the application. Remy's computer-generated algorithms can
achieve higher performance and greater fairness than some
sophisticated human-designed schemes, including ones that put
intelligence inside the network.

On a simulated 15~Mbps fixed-rate link with eight senders contending and
an RTT of 150~ms, a computer-generated congestion-control algorithm
achieved the following improvements in median throughput and
reductions in median queueing delay over these existing protocols:

\begin{center}

\begin{tabular}{|l|c|c|}
\hline
Protocol & Median speedup & Median delay reduction \\
\hline
\hline
Compound & $2.1\times$ & $2.7\times$ \\
NewReno & $2.6\times$ & $2.2\times$ \\
Cubic & $1.7\times$ & $3.4\times$ \\
Vegas & $3.1\times$ & $1.2\times$ \\
\hline
Cubic/sfqCoDel & $1.4\times$ & $7.8\times$ \\
XCP & $1.4\times$ & $4.3\times$ \\
\hline
\end{tabular}

{\footnotesize Adapted from \S\ref{sec:remyresults}.}

\end{center}

Once we had a computerized protocol designer, we explored using it as
a tool to gain understanding about the problem of congestion control.
How easy is it to ``learn'' a network protocol to achieve desired
goals, given a necessarily imperfect model of the networks where it
ultimately will be deployed? Is there a tradeoff between the
performance of an algorithm now---if optimized for a very specific
network---versus the protocol's ability to perform adequately as the
network evolves? What is the cost of compatibility with existing TCP
congestion-control mechanisms?

Our research into these questions is reported in
Chapter~\ref{chap:learnability}. We found weak evidence of a tradeoff
between the breadth of the operating range of a computer-generated
protocol and its performance, but also that a single
computer-generated protocol was able to outperform existing schemes
over a thousand-fold range of link rates.

We also measured the benefits and costs of adding ``TCP-awareness'' to
a computer-generated congestion-control protocol protocol. Including
the possibility of TCP cross-traffic in the network model improved
performance when TCP cross traffic was actually present---but hurt
performance when TCP cross traffic was absent, compared with a
``TCP-naive'' protocol designed only to interoperate with itself.

\begin{comment}
\subsection{Mosh}

\textbf{Mosh} (mobile shell)~\cite{mosh} is a replacement for SSH
(secure shell) in which the client and server each explicitly model
the dynamic contents of the text terminal. Unlike SSH, which transmits all
output from an application through a reliable TCP connection and
doesn't attempt to interpret the stream, Mosh's modeling allows it
to choose which data to send to the client when, in order to most
efficiently satisfy an explicit objective: update the client screen to
the current contents of the text terminal as efficiently as possible,
even across intermittent, roaming connectivity.

By having an explicit model of the user interface at both sides of the
connection, the Mosh client can predict the application's behavior in
response to user keystrokes and (if confident in its predictions)
speculatively display the result---generally an echoed keystroke or
cursor motion---before it has been confirmed by the server. In a
trace-based evaluation over 40 hours of real-world usage, Mosh was
able to successfully predict and immediately display the result of
\textbf{70\%} of user keystrokes.

Mosh is primarily a remote terminal application and is not further
discussed in this dissertation, but its design ideas---maintain an
explicit model at runtime, and pursue a stated objective on behalf of
the user---formed the basis of my subsequent work on
transport-protocol design.
\end{comment}