### 2014-05-21 concl.tex
Add discussion

\chapter{Discussion}

The practical impact of Sprout and Remy so far remains
preliminary. There have been several experiments on Sprout conducted
by companies interested in using it as a control algorithm for
videoconferencing or screen-sharing software. Generally speaking, I understand these
experiments have confirmed the ones in this thesis, under
the assumption that Sprout's request for bytes from the application
can always be satisfied.

However, in practice, coupling Sprout to an actual video coder has
presented several new challenges. Real video coders generally do not
have the agility to produce frames of exactly the requested coded
size, immediately on demand. Today's videoconferencing applications
typically allow the video coder to ``drive'' the application, coding and
sending video at a particular nominal bit rate until the transport
receives definitive evidence of congestion. Flipping that around---so
that Sprout's stochastic forecasts can ``drive'' the video
coder---may require changes to the video coder to be able to vary
the coded size of frames more rapidly.

As for Remy, much remains unknown about the capabilities and limits of
computer-generated algorithms, much less decentralized algorithms that
cooperate indirectly across a network to achieve a common
goal. Although the RemyCCs appear to work well on networks whose
parameters fall within or near the limits of what they were prepared
for---even beating in-network schemes at their own game and even
when the design range spans three orders of magnitude---our ability to reason about \emph{why} they
work is limited.

We can observe apparent limit cycles with the Ratatouille tool in
simulation and make some statements about the behavior of RemyCCs on
new kinds of simulated networks, but we do not have a principled
method to predict a RemyCC's performance on real-world heterogeneous
networks with unknown link parameters, topology, and cross-traffic.
Indeed, we have yet to run RemyCCs on real-world networks at all.

Along with the rigor of the computer-generated approach come new
challenges for protocol designers, who must figure out how to encode
their intuitions about the problem into an explicit statement of
assumptions and an objective. How should a protocol designer come up
with their model of the uncertain network? Our learnability
experiments in Chapter~\ref{chap:learnability} suggest that, even on
networks with more-complex structure, it may be sufficient for a
RemyCC to have been designed for the link rate and degree of
multiplexing of its \emph{bottleneck} link only. Much work needs to be
done, however, before we can reliably extrapolate from results on a
two-bottleneck parking-lot topology to more-complex networks. It may
also be possible for a RemyCC to ``learn'' different congestion
policies for different remote IP addresses, but the cross-interaction
between such adaptive RemyCCs---who might be executing different
policies at runtime---will need to be studied.

It was a surprising result that computer-generated end-to-end RemyCCs
were able to outperform some of the most sophisticated human-designed
in-network schemes, including Cubic-over-sfqCoDel and XCP. On the one
hand, this result is encouraging---it indicates that end-to-end
algorithms are more powerful than was previously understood,
perhaps making it possible to preserve a ``dumb'' network with smart endpoints.

In general, though, an in-network algorithm that can run on the
bottleneck gateway and enforce fairness among flows (as sfqCoDel and
XCP both do) is strictly more powerful than a decentralized algorithm
that runs purely end-to-end. My colleagues and I have
argued~\cite{sivaraman2013no} that, like endpoint congestion-control
schemes, in-network algorithms will need to keep evolving with time. I
expect that if similar methods to Remy are applied to the development
of in-network algorithms, they will be able to reclaim their
superiority over the best end-to-end algorithms.